---
title: "hw02-tsmith93"
author: "Thomas Smith"
date: '2018-09-19'
output: github_document
---
Load gapminder and tidyverse
```{r}
library(gapminder)
library(tidyverse)
```

What type of file?
```{r}
class(gapminder)
```
The function class (above) shows this is a tibble data frame, tibble, and a data frame

You can fin out the number of observations/rows and variable/collumns, respectively: 
```{r}
dim(gapminder)
```
This shows there are 6 collumns/variables and 1704 rows/observations

Alternatively, you can use individual functions to identify either number of rows:
```{r}
nrow(gapminder)
```

Or collumns:
```{r}
ncol(gapminder)
```

You can also figure out what is the type of each variables individually:
```{r}
class(gapminder$country)
```

Or for all of them!
```{r}
lapply(gapminder, class)
```

You can also look at the range of variables, whether they be categorical:
```{r}
levels(gapminder$country)
```

Or quantitative:
```{r}
range(gapminder$lifeExp)
```

Here is a summary of the data, including distribution:
```{r}
summary(gapminder)
```

Specific summary statistics can also be shown, such as the mean of each variable:
```{r}
lapply(gapminder, mean)
```

Next we can plot some of these variables. To start off, we will do a scatterplot of two variables. But before then, it is always good practice to make a grammar component table:

| Grammar Component     | Specification |
|-----------------------|---------------|
| __data__              | `gapminder` |
| __aesthetic mapping__ | `x`and `y` |
| __geometric object__  | point |
| scale                 | linear  |
| statistical transform | none |
| coordinate system     | rectangular |
| facetting             | none |

```{r}
ggplot(gapminder, aes(x=lifeExp, y=pop)) + 
    geom_point()
```

We can also present data for one quantitative variable using plots such as histograms:

| Grammar Component     | Specification |
|-----------------------|---------------|
| __data__              | `gapminder` |
| __aesthetic mapping__ |  `x` |
| __geometric object__  |  histogram|
| scale                 |  linear |
| statistical transform |  none |

```{r}
ggplot(gapminder, aes(gdpPercap)) +
    geom_histogram()
```

Finally, we can plot quantitative and categorical data together. A popular choice for this is a boxplot:

| Grammar Component     | Specification |
|-----------------------|---------------|
| __data__              | `gapminder` |
| __aesthetic mapping__ |  `x` and `y`|
| __geometric object__  |  boxplot |
| scale                 |  log-y scale |
| statistical transform |  5-number summary |

```{r}
a <- ggplot(gapminder, aes(continent, lifeExp)) +
  scale_y_log10()
a
```

We are also able to filter our data in order to plot data in a more precise way:
```{r}
select(gapminder, country, year, lifeExp)
```

Alternatively, filter:
```{r}
filter(gapminder, lifeExp>34)
```

Another way is to use piping:
```{r}
gapminder %>% 
  filter(lifeExp<30)
```

Now, why don't we apply filters to `ggplot()`!

| Grammar Component     | Specification |
|-----------------------|---------------|
| __data__              | `gapminder` |
| __aesthetic mapping__ |  |
| __geometric object__  |  |
| scale                 |  |
| statistical transform |  |

```{r}
gapminder %>% 
 filter(country == "India") %>% 
 ggplot(aes(year, pop)) +
 geom_line() +
 geom_point()
```

Or we can even plot the same information for multiple countries at the same time:

```{r}
c <- ggplot(gapminder, aes(year, lifeExp))
c + geom_line()
c + geom_line(aes(group=country), alpha=0.2)
```

Now for some extra fun, lets evaluate this data line where the author was hoping to get data only from Afghanistan and Rwanda.

```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```

It appears they did so correctly!

TA DA, done... almost...



