---
title: "hw02-tsmith93"
author: "Thomas Smith"
date: '2018-09-19'
output: html_document
---
Load gapminder and tidyverse
```{r}
library(gapminder)
library(tidyverse)
```

What type of file?
```{r}
class(gapminder)
```
The function class (above) shows this is a tibble data frame, tibble, and a data frame

You can fin out the number of observations/rows and variable/collumns, respectively: 
```{r}
dim(gapminder)
```
This shows there are 6 collumns/variables and 1704 rows/observations

Alternatively, you can use individual functions to identify either number of rows:
```{r}
nrow(gapminder)
```

Or collumns:
```{r}
ncol(gapminder)
```

You can also figure out what is the type of each variables individually:
```{r}
class(gapminder$country)
```

Or for all of them!
```{r}
lapply(gapminder, class)
```

You can also look at the range of variables, whether they be categorical:
```{r}
levels(gapminder$country)
```

Or quantitative:
```{r}
range(gapminder$lifeExp)
```

Here is a summary of the data, including distribution:
```{r}
summary(gapminder)
```

Specific summary statistics can also be shown, such as the mean of each variable:
```{r}
lapply(gapminder, mean)
```

We are also able to filter our data in order to plot data in a more precise way:
```{r}
select(gapminder, country, year, lifeExp)
```

Alternatively, filter:
```{r}
filter(gapminder, lifeExp>34)
```

Another way is to use piping:
```{r}
gapminder %>% 
  filter(lifeExp<30)
```

TA DA, done... almost...



